---
title: "Lab 3"
author: "Majood Haddad"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    theme: tactile
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Tasks

## Task 1
```{r}
getwd()
```

## Task 2
```{r}
spruce <- read.csv("SPRUCE.csv")
head(spruce)
```

## Task 3
Part 1: Scatter Plot

```{r}
plot(Height~BHDiameter, bg="blue", pch=21,cex=1.2, ylim = c(0,1.1*max(Height)), xlim =c(0,1.1*max(BHDiameter)), main="Height of Spruce trees vs. Breast Height Diameter", data=spruce)
```

It appears that there is no straight line relationship. It seems to be more of an exponential relationship than a straight line relationship.

Part 2:
```{r}
library(s20x)
layout(matrix(1:3,nr=3,nc=1))
trendscatter(Height~BHDiameter,f=0.5, data=spruce)
trendscatter(Height~BHDiameter,f=0.6, data=spruce)
trendscatter(Height~BHDiameter,f=0.7, data=spruce)
```

Part 3:
Linear Model Object
```{r}
spruce.lm = lm(Height~BHDiameter, data = spruce)
summary(spruce.lm)
```

```{r}
plot(Height~BHDiameter, data = spruce)
abline(spruce.lm)
```

Looking at the graph, the straight line is not appropriate as it does not represent the relationship between Height and BHDiameter in a strong way, and that can be seen from the R-squared value obtained from the summary function. However, when considering the curve line, it can be seen that the line includes more points and has a stronger representation of the relationship between the Height and BHDiameter variables.

## Task 4
```{r}

```

## Task 5
```{r}
summary(spruce.lm)
```

The value of the slope is 0.48147

The value of the intercept is 9.14684

The equation of the fitted line is y = 0.48147x + 9.14684

Predicting the height of the spruce:
```{r}
new_spruce <- data.frame(
  BHDiameter = c(15,18,20)
)
predict(spruce.lm, newdata = new_spruce)
```

Therefore, the predictions for the height of the spruce when the diameter is 15,18 and 20 are 16.36895, 17.81338 and 18.77632 respectively.

## Task 6
```{r}
library(ggplot2)
s <- ggplot(spruce.lm, aes(x=BHDiameter, y=Height, col = BHDiameter)) + geom_point() + geom_line() + geom_smooth() + ggtitle("Height vs. Diameter")
s
```

## Task 7
